---
title: "Unfair Coins"
date: "2/20/2020"
output: html_document
---

Initial question:
A jar holds 1000 coins. Out of all of the coins, 999 are fair and one is double-sided withtwo heads. Picking a coin at random, you toss the coin ten times.
Given that you see 10 heads, what is the probability that the coin is double headed and the probability that the next toss of the coin is also a head?


Lets take a look at this question and then play around with some of the given assumptions.
First, lets breakdown the original:
```{r}
NumberOfFlips = 10
NumberOfFairCoins = 999
NumberOfUnfairCoins = 1
FairOdds = .5
UnfairOdds = 1

# Probability of a fair or unfair coin
ProbabilityFair = NumberOfFairCoins / (NumberOfFairCoins + NumberOfUnfairCoins) # 999/1000
ProbabilityUnfair = NumberOfUnfairCoins / (NumberOfFairCoins + NumberOfUnfairCoins) # 1/1000
```

Now, to figured out the 11th flip, we'll use Bayes Theorem, or P(A|B) = P(B|A) * P(A) / P(B). Written out for this question, you would ask: what is the probability of Heads, given a fair coin.  That is equal to the probability of Heads given the unfair coin over multiplied against the probability of a fair coin over the probability of an unfair coin.

To start you need to ask what is the probability of a fair or unfair coin given 10 flips all heads:

```{r fair}
#P(Fair|10H) = P(10H | Fair) * P(Fair) / P(10H)

Probability10HGivenFair = FairOdds^NumberOfFlips
Probability10H = FairOdds^NumberOfFlips * ProbabilityFair + UnfairOdds^NumberOfFlips * ProbabilityUnfair

#ProbFairCoinGiven10Flips = 
ProbFairCoinGiven10Flips = Probability10HGivenFair * ProbabilityFair / Probability10H
ProbFairCoinGiven10Flips
```

```{r unfair}
Probability10HGivenUnfair = UnfairOdds^NumberOfFlips
Probability10H = FairOdds^NumberOfFlips * ProbabilityFair + UnfairOdds^NumberOfFlips * ProbabilityUnfair

ProbUnFairCoinGiven10Flips = Probability10HGivenUnfair * ProbabilityUnfair / Probability10H

ProbUnFairCoinGiven10Flips

```

Quick check that the the two probabilities add up to 1.

```{r check}
1 == ProbUnFairCoinGiven10Flips+ProbFairCoinGiven10Flips
```

So odds we have a fair coin is 49.38211% and the odds we have the unfair coin is 50.61789%.

So the odds the next flip will be heads would be the odds it heads given a fair coin (.5 * .4938) + odds if its unfair (1 * .5062) or:
``` {r nextToss}
FairOdds * ProbFairCoinGiven10Flips + UnfairOdds * ProbUnFairCoinGiven10Flips
```


Now, lets take a look at what would happen if we had more or less flips before estimating what the next flip would be.

We'll take the above work, but we'll make a data table for tracking results, and use a loop to run through flips.  And then plot the probability that the nth next flip with be heads.

```{r changeFlips}
FlipTable = data.table::data.table(NumberOfFlips = 1:20, ProbHead = 1)
for (i in 1:20){
  #Set up variable for table
  NumberOfFlips = i
  
  # Fit above work into a single formula
  FlipTable$ProbHead[i] = FairOdds *  (FairOdds^NumberOfFlips*ProbabilityFair/
      (FairOdds^NumberOfFlips * ProbabilityFair + UnfairOdds^NumberOfFlips * ProbabilityUnfair)) +
    UnfairOdds * (UnfairOdds^NumberOfFlips*ProbabilityUnfair / 
      (FairOdds^NumberOfFlips * ProbabilityFair + UnfairOdds^NumberOfFlips * ProbabilityUnfair))
}

plot(x = FlipTable$NumberOfFlips, y=FlipTable$ProbHead, type = "l")
```

We can see that around the 15th flip, its pretty much gaurunteed that the next flip will be heads and you almost certainly have the unfair coin.

So lets also see what happens if we change the number of unfair coins, what thes the probability of the 11th flip being heads if we add more unfair coins.

``` {r changeCoins}
CoinTable = data.table::data.table(NumberOfCoins = 1:50, ProbHead = 1)
for (i in 1:50){
  #Set up variable for table and reset number of flips
  NumberOfFairCoins = 1000-i
  NumberOfUnfairCoins = i
  NumberOfFlips = 10
  ProbabilityFair = NumberOfFairCoins / (NumberOfFairCoins + NumberOfUnfairCoins) # (1000-i)/1000
  ProbabilityUnfair = NumberOfUnfairCoins / (NumberOfFairCoins + NumberOfUnfairCoins) # i/1000
  
  #Combine above work into a single formula
  CoinTable$ProbHead[i] = FairOdds *  (FairOdds^NumberOfFlips*ProbabilityFair/
      (FairOdds^NumberOfFlips * ProbabilityFair + UnfairOdds^NumberOfFlips * ProbabilityUnfair)) +
    UnfairOdds * (UnfairOdds^NumberOfFlips*ProbabilityUnfair / 
      (FairOdds^NumberOfFlips * ProbabilityFair + UnfairOdds^NumberOfFlips * ProbabilityUnfair))
}

plot(x = CoinTable$NumberOfCoins, y=CoinTable$ProbHead, type = "l")


```

By adding more coins, we have an explosive increase that is over 95% by the 9th unfair coin.





